---
title: "Indian Liver Patient Records"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(dplyr)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")
if(!require(cowplot)) install.packages("cowplot", repos = "http://cran.us.r-project.org")
if(!require(grid)) install.packages("grid", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("gridExtra", repos = "http://cran.us.r-project.org")
if(!require(ggExtra)) install.packages("ggExtra", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")
if(!require(PerformanceAnalytics)) install.packages("PerformanceAnalytics", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(e1071)) install.packages("e1071", repos = "http://cran.us.r-project.org")
if(!require(randomForest)) install.packages("randomForest", repos = "http://cran.us.r-project.org")
if(!require(leaps)) install.packages("leaps", repos = "http://cran.us.r-project.org")

library(ggplot2)
library(dplyr)
library(PerformanceAnalytics)
library(e1071)
library(randomForest)
library(leaps)
library(knitr)

#Download data set "Indian Liver Patient Records""
df_IndianPatient<- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/00225/Indian%20Liver%20Patient%20Dataset%20(ILPD).csv", header = TRUE)

colnames(df_IndianPatient)<-c("Age", "Sex", "TB", "DB", "Alkphos", "Alamine", 
    "Aspartate", "TP", "ALB", "A_G_Ratio", "Disease")

df_IndianPatient<-na.omit(df_IndianPatient)

df_IndianPatient[which(df_IndianPatient$Disease==2),"Disease"]<-0

```

## Introduction

We download and use the data set ILPD (Indian Liver Patient Dataset) from web site kaggle (https://archive.ics.uci.edu/ml/datasets/ILPD+(Indian+Liver+Patient+Dataset))
This data set contients the Indian patients that have a liver diseas and not liver, and our goal is predict status of liver patients or not

## Data Analisys

The data set is composed from 11 variable:
\newline 1.  Age : Age of the patient
\newline 2.  Sex: Gender of the patient
\newline 3.  TB: Total Bilirubin
\newline 4.  DB: Direct Bilirubin
\newline 5.  Alkphos: Alkaline Phosphotase
\newline 6.  Alamine: Alamine Aminotransferase
\newline 7.  Aspartate: Aspartate Aminotransferase
\newline 8.  TP: Total Protiens
\newline 9.  ALB: Albumin
\newline 10. A_G_Ratio: Ratio Albumin and Globulin Ratio
\newline 11. Diesease: Selector field used to split the data into two sets (labeled by the experts)

we see the first rows of data set:
```{r head IDLR}
head(df_IndianPatient)
```



Most variabiles are type integer or numerical but only the variable "Sex" is a factor

```{r str IDLR}
str(df_IndianPatient)
```

We see the summary of data set and their distribution

```{r summary IDLR}
summary(df_IndianPatient)
```



## Data Visualization

The type of Sex in the Indian Patients are more of type Male than Female


```{r Sex, echo=FALSE}
df_IndianPatient %>% ggplot(aes(x=Sex,fill= Sex,group=Sex)) + geom_bar()  + 
  labs(title="Bar plot for type of Sex") 
```

The Male Patients are 441 and the Female are 141 

We see the distribution of Age in the histogram, that is similar at the gaussian.

```{r Age, echo=FALSE}
df_IndianPatient %>% ggplot(aes(x=Age)) +
                 geom_histogram(breaks=seq(min(df_IndianPatient$Age), max(df_IndianPatient$Age), by = 3), 
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram for Age") +
  labs(x="Age", y="Count") + 
  xlim(c(min(df_IndianPatient$Age),max(df_IndianPatient$Age))) + 
  ylim(c(0,(as.numeric(which.max(table(df_IndianPatient$Age)))+5)))

```

and the age for type of sex

```{r Age4Sex, echo=FALSE}
df_ILPD_Male<-df_IndianPatient %>% filter(Sex=="Male") %>% ggplot(aes(x=Age)) +
                 geom_histogram(breaks=seq(min(df_IndianPatient$Age), max(df_IndianPatient$Age), by = 3), 
                 col="Blue", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram for Men Age") +
  labs(x="Age", y="Count") + 
  xlim(c(min(df_IndianPatient$Age),max(df_IndianPatient$Age))) + 
  ylim(c(0,(as.numeric(which.max(table(df_IndianPatient$Age)))+5)))

df_ILPD_Female<-df_IndianPatient %>% filter(Sex=="Female") %>% ggplot(aes(x=Age)) +
                 geom_histogram(breaks=seq(min(df_IndianPatient$Age), max(df_IndianPatient$Age), by = 3), 
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram for Female Age") +
  labs(x="Age", y="Count") + 
  xlim(c(min(df_IndianPatient$Age),max(df_IndianPatient$Age))) + 
  ylim(c(0,(as.numeric(which.max(table(df_IndianPatient$Age)))+5)))

library(cowplot)
cowplot::plot_grid(df_ILPD_Male, df_ILPD_Female, labels = "AUTO")

```



The Disease for type of Sex have this distribution:


```{r tab Disease4Sex, echo=FALSE}
table(df_IndianPatient$Disease,df_IndianPatient$Sex)

```


```{r Disease, echo=FALSE}
df_IndianPatient$Disease<-as.factor(df_IndianPatient$Disease)
df_IndianPatient %>% ggplot(aes(x=Sex,fill= Disease,group=Disease)) + geom_bar()
```

There are in the data set most cases with Disease Liver Patient

The Bilorubin is define as yellow compound that occurs in the normal catabolic pathway that breaks down heme in vertebrates. This catabolism is a necessary process in the body's clearance of waste products that arise from the destruction of aged or abnormal red blood cell (for other information to Bilirubin to see https://en.wikipedia.org/wiki/Bilirubin)

We rappresent the relation beetwen "Total Bilurubin" and "Direct Bilirubin"

```{r RegressionBilorubin, echo=FALSE}
library(ggExtra)
plot_center = ggplot(df_IndianPatient, aes(x=TB,y=DB)) + 
  geom_point() +
  geom_smooth(method="lm")

# default: type="density"
ggMarginal(plot_center, type="histogram")
```

where there is a linear correlation 

After we see  the Alkphos, that is a homodimeric protein enzyme of 86 kilodaltons.
(For other information to Alkphos to see https://en.wikipedia.org/wiki/Alkaline_phosphatase)


```{r Alkphos, echo=FALSE}
df_IndianPatient %>% ggplot(aes(x=Alkphos)) +
                 geom_histogram(breaks=seq(min(df_IndianPatient$Alkphos), max(df_IndianPatient$Alkphos), by = 3), 
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram for Alkphos") +
  labs(x="Alkphos", y="Count") + 
  xlim(c(min(df_IndianPatient$Alkphos),max(df_IndianPatient$Alkphos))) + 
  ylim(c(0,(as.numeric(which.max(table(df_IndianPatient$Alkphos))))))
```



We see the ralation beetwen Alamine Aminotransferase (was formerly called serum glutamate-pyruvate transaminase (SGPT) or serum glutamic-pyruvic transaminase (SGPT)) and Aspartate Aminotransferase ( is a pyridoxal phosphate (PLP)-dependent transaminase enzyme)


```{r Aminotransferase, echo=FALSE}
plot_center = ggplot(df_IndianPatient, aes(x=Alamine,y=Aspartate)) + 
  geom_point() +
  geom_smooth(method="lm")

# default: type="density"
ggMarginal(plot_center, type="histogram")
```

There is a linear correaltion beetwen two variable


Now we can see the relation beetwen all variable, in an unique plot

```{r CorrealtionIDLR, echo=FALSE}
df_IDLRnumeric<-df_IndianPatient[,-2]
df_IDLRnumeric[,1]<-as.numeric(df_IDLRnumeric[,1])
df_IDLRnumeric[,2]<-as.numeric(df_IDLRnumeric[,2])
df_IDLRnumeric[,3]<-as.numeric(df_IDLRnumeric[,3])
df_IDLRnumeric[,4]<-as.numeric(df_IDLRnumeric[,4])
df_IDLRnumeric[,5]<-as.numeric(df_IDLRnumeric[,5])
df_IDLRnumeric[,6]<-as.numeric(df_IDLRnumeric[,6])
df_IDLRnumeric[,7]<-as.numeric(df_IDLRnumeric[,7])
df_IDLRnumeric[,8]<-as.numeric(df_IDLRnumeric[,8])
df_IDLRnumeric[,9]<-as.numeric(df_IDLRnumeric[,9])
df_IDLRnumeric[,10]<-as.numeric(df_IDLRnumeric[,10])
chart.Correlation(as.data.frame(df_IDLRnumeric), histogram=TRUE, pch=19)
```

there are the following relations
\newline - "TB" and "DB"" 
\newline - "Alamine" and "Aspartate"
\newline - "TP" and "ALB"
\newline - "ALB" and "A_G_Ratio"

 that are all linear relation most significative
 
## Model of analisys

Now create our set data to predict the model

```{r CreateTestValidSet}
set.seed(7) # for reproducibility
test_index <- createDataPartition(y = df_IndianPatient$Disease, times = 1, p = 0.7, list = FALSE)
edx <- df_IndianPatient[-test_index,]
temp <- df_IndianPatient[test_index,]
```


## Logistic Regression

We applicate the logistic regression to predict a model, because the output variable assume value 0 and 1

```{r ModelPredict}
fit <- glm(Disease ~ Age + Sex + TB + DB + Alkphos + Alamine + Aspartate + 
    TP + ALB + A_G_Ratio, data = edx, family = binomial(link = "logit"))

summary(fit)
```

Calculate the accuracy of model, that is


```{r AccuracyLogistic}
Edx_Predictions <- data.frame(Probability = predict(fit, edx, type = "response"))
Edx_Predictions$Prediction <- ifelse(Edx_Predictions > 0.5, 1, 0)
Edx_Predictions$Disease <- edx$Disease
accuracy <- mean(Edx_Predictions$Disease == Edx_Predictions$Prediction, na.rm = TRUE)
tot_accuracy<-data.frame(Model="Logistic Regression",Value=accuracy)

accuracy


```


## Backward in Logistic Regression

Now we applicate the backward for select the best significative variabiles to rappresent the model.

```{r Backward, echo=FALSE}
regfit.full=regsubsets(Disease~.,edx, method ="backward")
summary(regfit.full)

```

The most significative variables for model are Age, DB, Alamine, Thart select the m

```{r BackwardLogistic}
backward.model <- glm(Disease ~ Age + DB + Alamine, data = edx, family = binomial(link = "logit"))
summary(backward.model)[7]
```

and we calculate the accuracy

```{r BackwardLogisticAccuracy}
predBw<-predict(backward.model,temp,type = "response")
predicted.BW <- as.numeric(ifelse(predBw > 0.5, 1, 0))
accuracyBackward<-mean(predicted.BW==temp$Disease,na.rm=TRUE)
tot_accuracy<-rbind(tot_accuracy,data.frame(Model="Backward Logistic Regression",Value=accuracyBackward))
accuracyBackward
```


## Random Forest

Predict a Model with approach Random Forest

```{r RF}
IDLR.rf=randomForest(Disease ~ ., data = edx,
              na.action=na.exclude)
 pred<-as.numeric(predict(IDLR.rf,temp))

```


and we have the accuracy model with this value:

```{r AccuracyRF, echo=FALSE}
predValid <- as.numeric(as.character(predict(IDLR.rf, temp, type = "class")))
# Checking classification accuracy
predValid<- ifelse(predValid > 0.5, 1, 0)
accuracyRF<-mean(as.numeric(predValid) == temp$Disease,na.rm=TRUE)  
tot_accuracy<-rbind(tot_accuracy,data.frame(Model="Random Forest",Value=accuracyRF))
accuracyRF
```

## Results
We see the results of accuracy of differents models applicate

```{r accuracyTot}
tot_accuracy %>% knitr::kable()
```


##Conclusion
We see the best model to predict the Diseas is the first model where we applicate the logistic regression.
This model have a good prediction for our data.


